---
title: "Day2_EPL"
author: "Bravo Kilo"
date: "2023-08-24"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Day 2 English Premiere League Dataset

## Loading Libraries

```{r message=FALSE, warning=FALSE, echo=FALSE}

library("dplyr")
library("tidyr")
library("readr")
library("purrr")
library("DT")

```

## Read the files in the data directory

```{r}

ssn1 <- read.csv("../../../Data/English Premier League_09_19/ssn-0910_csv.csv")
ssn2 <- read.csv("../../../Data/English Premier League_09_19/ssn-1011_csv.csv")
ssn3 <- read.csv("../../../Data/English Premier League_09_19/ssn-1112_csv.csv")
ssn4 <- read.csv("../../../Data/English Premier League_09_19/ssn-1213_csv.csv")
ssn5 <- read.csv("../../../Data/English Premier League_09_19/ssn-1314_csv.csv")
ssn6 <- read.csv("../../../Data/English Premier League_09_19/ssn-1415_csv.csv")
ssn7 <- read.csv("../../../Data/English Premier League_09_19/ssn-1516_csv.csv")
ssn8 <- read.csv("../../../Data/English Premier League_09_19/ssn-1617_csv.csv")
ssn9 <- read.csv("../../../Data/English Premier League_09_19/ssn-1718_csv.csv")
ssn10 <- read.csv("../../../Data/English Premier League_09_19/ssn-1819_csv.csv")

ssn1 %>% 
  head() %>% 
  knitr::kable()

```

## Merging the datasets

```{r}

epl <- merge(ssn1, ssn2, all = TRUE)
epl <- merge(epl, ssn3, all = TRUE)
epl <- merge(epl, ssn4, all = TRUE)
epl <- merge(epl, ssn5, all = TRUE)
epl <- merge(epl, ssn6, all = TRUE)
epl <- merge(epl, ssn7, all = TRUE)
epl <- merge(epl, ssn8, all = TRUE)
epl <- merge(epl, ssn9, all = TRUE)
epl <- merge(epl, ssn10, all = TRUE)

dim(epl)

```


## option 2 merge

```{r}

epl1 <- bind_rows(
  ssn1
,ssn2
,ssn3
,ssn4
,ssn5
,ssn6
,ssn7
,ssn8
,ssn9
,ssn10
)

dim(epl)

```

## Option 3

```{r}

epl2 <- list.files("../../../Data/English Premier League_09_19/", 
           full.names = TRUE, pattern = "*.csv") %>% 
  set_names() %>% 
  map_dfr(~read.csv(.), .id = "Source")

dim(epl2)

```

The number of games played at home for arsenal `r 2+3`

The number of games played at home by Arsenal `r sum(epl2$HomeTeam == "Arsenal")` and away games were `r sum(epl2$AwayTeam == "Arsenal")`. 

The total number of games played by Arsenal `r sum(epl2$HomeTeam == "Arsenal") + sum(epl2$AwayTeam == "Arsenal")`


Which team had the highest number of home goals?

```{r echo=FALSE}
HighestFTHG <-  epl2 %>% 
  group_by(HomeTeam) %>% 
  summarise("SumofHomeGoals" = sum(FTHG)) %>% 
  arrange(desc(SumofHomeGoals))
```


The team is `r HighestFTHG$HomeTeam[1]`



### Which team has the highest goals under the three categories total goals, home games and away games


```{r}

epl2_reduced <- epl2 %>% 
  select(Source, Date, HomeTeam, AwayTeam,
         FTHG, FTAG, FTR)

datatable(epl2_reduced)


```
#### Change the data from wide to long format

```{r}

epl2_reduced_long <- epl2_reduced %>% 
  pivot_longer(cols = c("HomeTeam", "AwayTeam"),
               names_to = "Home_Away", values_to = "Team")

dim(epl2_reduced_long)

```

```{r}

epl2_reduced_long <- epl2_reduced_long %>% 
  mutate(FT = if_else(Home_Away == "HomeTeam", FTHG, FTAG))


```

```{r}
epl2_reduced_long_sum <- epl2_reduced_long %>% 
  select(Home_Away, Team, FT) %>% 
  group_by(Home_Away, Team) %>% 
  mutate("SumOfGoals" = sum(FT)) %>% 
  ungroup() %>% 
  select(-FT) %>% 
  distinct()

datatable(epl2_reduced_long_sum)

```

```{r}

epl2_reduced_long_sum_t <- epl2_reduced_long_sum %>% 
  pivot_wider(names_from = "Home_Away", values_from = "SumOfGoals")

datatable(epl2_reduced_long_sum_t,
   filter = "top")

```

